---
title: "Tinnitus Study"
author: "Beatriz Vela"
date: "2023-08-24"
output: html_document
---

This is a project completed during Fall 2021 for my Statistical Learning class as part of my MS in Applied Statistics and Data Science. At first, the report of this project was done using Microsoft Word. However, I've now transitioned into RMarkdown specifically for my portfolio. This change lets me showcase the project in a more engaging and interactive way.

## Introduction 

Tinnitus can be described as a sensation of hearing ringing or buzzing sounds even in the absence of external sounds. Often, tinnitus interferes with daily tasks resulting in psychological distress. To help tinnitus patients cope with the psychological effects of tinnitus, an internet based cognitive behavioral therapy (ICBT) was developed. For this case study, a data set containing the information of 142 subjects experiencing Tinnitus has been provided. Information provided by the data set are: pre-study and post-study tinnitus scores (Pre_TFI and Post_TFI scores), clinical information such as a hearing survey score (HHI_Score), Generalized Anxiety Disorder score (GAD), depression sum (PHQ), Insomnia (ISI), satisfaction with life (SWLS), cognitive failures (CFQ), duration of tinnitus, and demographic information such as 
Gender and Age.

## Methodology
In R software the following libraries were used:

- `tidyverse`: to remove variables as needed
- `corrplot`: for correlation matrix
- `caret`: for data partition
- `leaps`, `olsrr`, and `MASS`: for model selection
- `nortest`:  for Anderson-Darling normality test
- `splines` and `gam`: for Generalized Additive Model (GAM)

## Exploring the Data 

First, setting the working directory and reading the data: 
```{r include=FALSE}
setwd("C:/Users/betyv/Documents/UTRGV/Fall 2021/Statistical Learning/Case Studies/Case Study 1")
getwd()
```

A working directory with the `.csv` file was set in our computer. Then, to read the data in R we use the `read.csv` command. 

To visualize the file and its predictors, we call for `head()`, and the see the dimensions of our file we use `dim()`. Here, we can see that the file contains 142 rows and 14 columns. 

```{r}
###Read the data 
casestudy1 <- read.csv(file = 'CaseStudy1.csv')
head(casestudy1)
dim(casestudy1)

```


